---
title: "ANOVA for Journalists: Analyzing College Admissions Test Scores"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Introduction

In this exercise, we'll explore a common journalism question: *Are there significant differences in standardized test scores across different types of colleges?*

We'll use ANOVA (Analysis of Variance) to determine if the observed differences in SAT scores across three college types (Public, Private, and Community College) represent real patterns or just random variation.

## Task 1: Load Libraries and Create Sample Data

```{r}
library(tidyverse)
```

```{r create-data}
# Set seed for reproducibility
set.seed(123)

# Create simulated SAT scores for admitted students across three college types
# Scale is 400-1600 for total SAT
public_college <- rnorm(100, mean = 1350, sd = 120)
private_college <- rnorm(100, mean = 1450, sd = 100)
community_college <- rnorm(100, mean = 1050, sd = 150)

# Combine into a single dataframe
admissions_data <- data.frame(
  college_type = factor(rep(c("Public", "Private", "Community"), each = 100)),
  sat_score = c(public_college, private_college, community_college)
)

# Ensure SAT scores are within realistic range (400-1600)
admissions_data$sat_score <- pmin(pmax(admissions_data$sat_score, 400), 1600)

# Quick look at the data
head(admissions_data)
```

## Task 2: Calculate Basic Statistics

```{r summary-stats}
# Summary statistics by college type
admission_summary <- admissions_data |>
  group_by(college_type) |>
  summarize(
    mean_score = mean(sat_score),
    median_score = median(sat_score),
    sd_score = sd(sat_score),
    min_score = min(sat_score),
    max_score = max(sat_score)
  )

print(admission_summary)
```

**Task 3** Describe the data

Write a couple of sentences describing the differences in SAT scores between types of colleges, using the descriptive statistics we just generated.

The data shows that students admitted private colleges have higher SAT scores than those admitted to public and community colleges, which has lower scores. The higher SAT scores show a lower standard deviation, while students admitted to community college's scores have a greater standard deviation. Public colleges scores and standard deviation is in the middle.

## Task 4: Visualize the Data

Let's create visualizations to see the patterns:

```{r visualize-distributions}
# Create density plot showing overlapping distributions
ggplot(admissions_data, aes(x = sat_score, fill = college_type)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "SAT Score Distributions by College Type",
    subtitle = "Are these differences statistically significant?",
    x = "SAT Score (400-1600 scale)",
    y = "Density",
    fill = "College Type"
  ) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1")
```

```{r boxplot}
# Create boxplot for another view
ggplot(admissions_data, aes(x = college_type, y = sat_score, fill = college_type)) +
  geom_boxplot() +
  labs(
    title = "SAT Scores by College Type",
    x = "College Type",
    y = "SAT Score"
  ) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1")
```

## Task 5: Perform ANOVA Test

```{r anova-test}
# Run one-way ANOVA
anova_result <- aov(sat_score ~ college_type, data = admissions_data)
summary(anova_result)
```

The F-value is huge (284.3). The F-value measures the ratio of between-group variance to within-group variance, and this high value suggests the differences between college types far outweigh the variation within each type. The p-value is very low, indicating the differences between college types are statistically significant at any reasonable significance level. The three asterisks (***) highlight that this is significant at the 0.001 level. So we can say that with confidence that there are real differences in SAT scores between the three types of colleges.

## Task 5: Compare the differences between specific pairs

```{r post-hoc}
# Tukey's HSD test to see which specific pairs of college types differ
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)
```

The Tukey multiple comparisons test shows highly significant differences in SAT scores between all college types:

* Private vs. Community: Private colleges have SAT scores that are 366.96 points higher than community colleges on average (p < 0.0001)
* Public vs. Community: Public colleges have SAT scores that are 292.55 points higher than community colleges on average (p < 0.0001)
* Public vs. Private: Public colleges have SAT scores that are 74.40 points lower than private colleges on average (p = 0.000021)

All comparisons show extremely small p-values, indicating these differences are statistically significant at any conventional significance level. The 95% confidence intervals (shown in the "lwr" and "upr" columns) do not include zero for any comparison, further confirming these differences are statistically significant.

## Task 6: Report the Findings

Based on our analysis, we can now answer:

1. Is there a significant difference in SAT scores across college types?
2. Which specific college types differ from each other?
3. How large are these differences?

Private college students scored on average 366.96 points higher than community college students and 74.4 points higher than public college students. Each comparison has a p-value of .001, allowing us to say that differences are statistically significant. In addition to the low p-value, the ANOVA test showed an F-value of 289.3, indicating that the variation in SAT sores between colleges is much grater than the variation within each group.


Answer those questions, providing evidence for your answers, and then write a headline and a lead paragraph reporting these findings using active voice and focusing on the differences across types.

Headline: SAT scores vary widely across U.S. college types. Here's what the data says

Lede: Where you go to college is strongly linked to SAT scores. A new analysis showed that students admitted to private colleges score significantly higher on the SAT than those at public and community colleges. The data shows a 370-point gap between private and community college students and over a 70-point gap between private and public college students.

