---
title: "Sampling"
---

Follow along with the demo to learn how to sample data from a population using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```


### Task 2: Load 1,000 rows of data with the ages of people from 18-80

```{r}

set.seed(42) #Setting seed for reproducibility
population <- tibble(
  age = sample(18:80, 1000, replace = TRUE)
)

#View the first few rows
head(population)

```


### Task 3: Sample 100 rows of data from the dataset

```{r}

sample_100 <- population |> 
  sample_n(100)

sample_50 <- population |> 
  sample_n(50)

sample_10 <- population |> 
  sample_n(10)

```


### Task 4: Calculate the mean age and standard deviation of the sample

```{r}

sample_100 |> summarize(mean = mean(age), sd = sd(age))

sample_50 |> summarize(mean = mean(age), sd = sd(age))

sample_10 |> summarize(mean = mean(age), sd = sd(age))


```


### Task 5: Calculate the mean age and standard deviation of the population

```{r}
population |> summarize(mean = mean(age), sd = sd(age))

```


### Task 6: Compare the sample and population statistics

Compare the means and standard deviations. Are they similar? Go back and re-run the sampling and statistics calculations with a smaller sample size. How do the results change?

The mean and standard deviation for the population of 1000 is fairly similar to if we looked at a sample of 50 and 100. It starts to deviate once you get to a population of 10, where the mean dropped from 49.873 to 41.2. The results are pretty different when you consider that the sample of 50, 100 and 1000, all remained in the 48-50 range.

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
