---
title: "Identifying Schools with Biggest Differences in Math Test Scores, Part 2"
author: Emely Miranda-Aguilar
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
options(scipen = 999)
```

### Introduction

We'll do what we did in class, but with a twist: instead of Montgomery County, Pa., you'll be working with data from Montgomery County, Md. Same basic idea, though: linear regression looking at free and reduced-price meals and math test scores.

Just as in the exercise, you will create a scatterplot to visualize the relationship between the FARMS (Free and Reduced-price Meal System) percentage and the percentage of students who met or exceeded math expectations. You'll then use linear regression to identify the schools that show the biggest differences between their actual and expected math test scores and evaluate the results of your model.

### Loading the Data

First, let's load the data into a data frame using the read_csv() function from the tidyverse library.

```{r}
schools <- read_csv("https://raw.githubusercontent.com/dwillis/jour405_files/refs/heads/main/montgomery_md.csv")
```
### Task: Create a Scatterplot and Describe the Results (5 points)

First, let's create a scatterplot to visualize the relationship between the FARMS percentage and the percentage of students who met or exceeded math expectations. We'll use ggplot2 to create the plot and add a linear regression line.

```{r}
schools |>
  ggplot(aes(x = `FARMS Pct`, y = `% Met or Exceeded Math Expectations`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "FARMS Percentage", y = "% Met or Exceeded Math Expectations", title = "Scatterplot of Math Expectations vs FARMS Percentage")
```
Based on what you see here, what's the relationship between FARMS percentage and math test scores? Compare this scatterplot to the one you made in the class activity.

The line shows a negative relationship. The points on the graph are closer to the line than the activity from Tuesday where there was no clear trend. Based off of the scatterplot, higher FARMS percentage is associated with lower percentage of math scores meeting/exceeding expectations.


### Task: Linear Regression Model (7 points)

Next, you'll perform a linear regression analysis using the lm() function. We'll set the percentage of students who met or exceeded math expectations as the dependent variable, and the FARMS percentage as the independent variable. In other words, if we know the FARMS percentage for a school, how reliably can we predict its math score?

```{r}
model <- lm(`% Met or Exceeded Math Expectations` ~ `FARMS Pct`, data = schools)

summary(model)
```

Provide the model's output for the three statistics mentioned in the class activity.

The p-value, the adjusted R-squared value and the coefficients.

Based on what you've learned so far, what does it tell us about the nature of the relationship between FARMS percentage and math scores? Is this model worth exploring further? Why or why not?

The p-value for this data is less than .05 (by a lot), this automatically tells us that the relationship is statistically significant. The adjusted R-square is .6734. This means that 67.34% of the FARMS percentage explains the difference between schools' math scores. The coefficient for the FARMS percentage is -0.595, causing a negative slope. On the scatterplot, it shows that higher FARMS percentage shows led to lower math scores.

### Task: Identifying Schools with Biggest Residuals (3 points)

Let's calculate and examine the residuals next. 

```{r}
schools$predicted <- predict(model)

# Calculate the residuals
schools <- schools |> 
  mutate(residual = `% Met or Exceeded Math Expectations` - predicted)

# Sort the data frame by the absolute value of the residuals in descending order
schools <- schools |>
  mutate(abs_residual = abs(residual)) |>
  arrange(desc(abs_residual))

```

Look at the absolute values of the residuals; what are the largest differences between the predicted result and the actual math scores? Pick two schools with the largest residuals. Find the zip codes where they are located and using the CensusReporter.org website, describe the most interesting demographic data from each zip code.

The largest differences are from Wayside Elementary and Stedwick Elementary. Wayside over-performed: the predicted score was 25.14%, but 81.6% of students met/exceeded expectations, resulting in an absolute residual of 56.46. Stedwick under-performed: the predicted score was 63.35%, but only 8.3% met/exceeded expectations, resulting in an absolute residual of 55.05.

Wayside's zip code: 20854
- 59% White and 22% Asian
- 5% Black and 9% Hispanic with at least 10% MoE (both) 
- 1% of children (age under 18) live in poverty with at least 10% MoE

Stedwick's zip code: 20886
- 35% Hispanic and 23% Black
- 23% White and 14% Asian with at least 10% MoE (both)
- 16% of children (age under 18) live in poverty with at least 10% MoE

## Task: Evaluate the model (5 points)

Describe what conclusions you'd draw from this model's results and how it would influence a story pitch about the school board member and math scores.

If the school board member is making the same claim as Tuesday's activity that schools with fewer students on FARMS do better on standardized math tests, then the model would show strong evidence of this claim.

The p-value is far below .05, showing a statistically significant relationship. The adjusted r-square show that 67.34% of the FARMS percentage explains the variaation in scores. Wayside is an example of a school in a zip code with a very low poverty rate and over-performed in math tests, unlike Stedwick, which under-performed by a lot (55.05 points). This would shape a story pitch by validating the overall trend but also pushing to explore what helps schools succeed or struggle beyond just FARMS percentages.
