---
title: "HW2 Prince George's County Crime"
---

In this assignment, we will be working with a dataset reporting the number of crimes by month in Prince George's County, Maryland, for every month in 2024. 

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv. You will create a dataframe called `pg_crime` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
pg_crime <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, are any of the categories more important than others? Why or why not? Do any of them make the job of summarizing the data more difficult? Why or why not?

There are many categories, with some being more specific than others depending on the type of incident that occurred. I would say the most important categories are "theft from auto" and "stolen vehicles" because these are the two crimes with the highest incidents, excluding "accident," which I believe refers to a car accident and is not necessarily a crime. The "accident with impound" category also has high number of incidents. I do not think it will be difficult to summarize the data. In general, the data is pretty straightforward, especially because the categories are specific with multiple of the same ones with just different circumstances. The only parts that might cause an issue is that some categories say "other" which can mean multiple things and residential robberies missing data for 2 months. Also some categories might overlap.

### Task 4: Calculating Averages

Calculate the total number of incidents per month for each category using `mutate()`, calling the new column `total_incidents`. Then calculate the average number of incidents per month for each category of crime, calling the new column `average_incidents`. Assign the result to a new dataframe called `pg_crime_averages`.

```{r}
total_incidents <- pg_crime |> mutate(total_incidents = (Jan+Feb+Mar+Apr+May+Jun+Jul+Aug+Sep+Oct+Nov+Dec))
pg_crime_average <- total_incidents |> mutate(average = (total_incidents/12))
```

### Task 5: Describe the Averages

Examine the new dataframe, `pg_crime_averages`, by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the average number of incidents per month for each category. Which categories have the highest average number of incidents per month? Which have the lowest? What does this tell you about the pattern of crime in Prince George's County? Are there categories that you have questions about? Why or why not?

The highest average number of incidents is "accident" with an average of 481.33 and an additional 320.75 "accident with impound". Other incidents that have a high average are "auto stolen" with 399.17 and "theft from auto" with 391.92. The lowest average number of incidents is homicide with an average of 5.83. "B&E other" also has a low average with 7.08. The pattern I realize is that PG county has a lot of incidents and crime involving vehicles. The only category I have a question about is "robbery residential" because we don't know the total or average. Why don't we have data for March and October?


### Task 6: Testing a Theory

One theory about crime is that it is more likely to occur in the summer months. Test this theory by creating three new columns using `mutate()` that calculate the percentage change compared to the average for each category of crime for the months of June, July, and August. Call the new columns `june_change`, `july_change`, and `august_change`. Assign the result to a new dataframe called `summer_crime_averages`.

```{r}
june_pct_change <- pg_crime_average |> mutate(june_change = (Jun-average)/average*100)
july_pct_change <- june_pct_change |> mutate(july_change = (Jul-average)/average*100) 
summer_crime_averages <- july_pct_change |> mutate(august_change = (Aug-average)/average*100)
```

### Task 7: Describe the Summer Crime Averages

Look at the `summer_crime_averages` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change in the average number of incidents per month for each category of crime for the months of June, July, and August. Do you see any patterns? What does this tell you about the theory that crime is more likely to occur in the summer months?

The data shows a lot of decrease in crime during the summer in general, but in the summer the month with the highest increase rate and lowest decrease rate is June. It has higher rates in seven incidents. Of the three summer months, July was second with high increase/low decrease rates in six incidents. August came third with high increase/low decrease in five incidents. A pattern I noticed is that the largest percent changes that showed an increase in crime occurred in July, specifically in "assault shooting", "assault weapon", "auto stolen & recovered", "robbery vehicle", "sex offense" and "theft from auto". This can mean that some violent crimes and incidents involving cars are more likely to occur in July than the other summer months.

### Task 8: What's the Story?

Write up a lede and nutgraf of a story focused on the crime data in Prince George's County. It could be from any of the dataframes you've created in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

Crimes involving vehicles are among the most frequent in Prince George's County.

Vehicle-related incidents had the highest crime rates in the county in 2024. According to Prince George's County 2024 crime report, nearly 5,000 cases of both "auto stolen" and "theft from auto" occurred. The county's report showed that last year an average of about 400 stolen vehicles a month, the highest of other incidents excluding accidents, whereas homicides averaged about 6 a month. Data showed that an additional 487 cases of stolen vehicles that were  recovered, which means that over 90% of stolen vehicles reported are not recovered. The report highlights the scale of vehicle-related crimes in the county compared to other crimes.

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
