---
title: "HW7_MandMs"
---

In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```


### Task 2: Load the M&Ms dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv

```{r}
mms <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv")
```


### Task 3: Sample 5 rows of data from the dataset, and then sample 10 rows, saving each to separate dataframes called sample5 and sample10. Be sure to use set.seed(1234) before you start.

```{r}
sample5 <- mms |> 
  sample_n(5)

sample10 <- mms |> 
  sample_n(10)
```


### Task 4: Calculate the mean and standard deviation of the two samples, using one of the colors.

```{r}
sample5 |> summarize(mean = mean(red), sd = sd(red))
sample10 |> summarize(mean = mean(red), sd = sd(red))
```


### Task 5: Calculate the mean and standard deviation of the original dataset, using the same color as Task 4

```{r}
mms |> summarize(mean = mean(red), sd = sd(red))
```


### Task 6: Make three histograms: one each for sample5, sample10 and the original dataset. Use a consistent binwidth.

```{r}
mms |> 
  ggplot() +
  geom_histogram(aes(x=red), fill = "red", color = "black", binwidth = 5) +
  geom_vline(aes(xintercept = mean(red)), color = "blue", linetype = "dashed") +
  geom_vline(aes(xintercept = mean(red) + sd(red)), color = "purple", linetype = "dashed") +
  geom_vline(aes(xintercept = mean(red) - sd(red)), color = "purple", linetype = "dashed")



sample5 |> 
  ggplot() +
  geom_histogram(aes(x=red), fill = "red", color = "black",  binwidth = 3) +
  geom_vline(aes(xintercept = mean(red)), color = "blue", linetype = "dashed") +
  geom_vline(aes(xintercept = mean(red) + sd(red)), color = "purple", linetype = "dashed") +
  geom_vline(aes(xintercept = mean(red) - sd(red)), color = "purple", linetype = "dashed")

sample10 |> 
  ggplot() +
  geom_histogram(aes(x=red), fill = "red", color = "black",  binwidth = 3) +
  geom_vline(aes(xintercept = mean(red)), color = "blue", linetype = "dashed") +
  geom_vline(aes(xintercept = mean(red) + sd(red)), color = "purple", linetype = "dashed") +
  geom_vline(aes(xintercept = mean(red) - sd(red)), color = "purple", linetype = "dashed")
```


### Task 7: Answer some questions

Compare the three sets of means, standard deviations and histograms. Which are more helpful to you in understanding the accuracy of the samples?

  I think the first one is more accurate because there is a lot more variety. The means for the other two (sample5 and sample10) are a little off from the original mean, so I don't think they are accurate.

What does this exercise tell you about the relationship between sample size and estimate accuracy?

I think the smaller the sample size the less accurate it is, and the larger the sample size the more accurate it is.

How does random sampling introduce variation into statistical estimates?

I don't know if this is what the question is asking, but I think random sampling introduces variation if the sample is being pulled from a population with a lot of a variety, or at least an equal amount of variety. However, I think that sometimes you might not end up with a truly diverse sample because it's random and the sample isn't strategically chosen to ensure that there is variety. You have to rely on chance.

If you repeated this process, would the samples always be identical? Why or why not?

I don't think so because the is random sampling, so each time it'll be random. I do think it's likely to be similar, but I wouldn't say identical.

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
