---
title: "final_project"
output: html_document
---

5/3/2025

USM HISPANIC ENROLLMENT:
```{r}
usm_hispanic_enrollment |>
  group_by(year) |>
  summarize(total = sum(enrollment)) |>
  ggplot(aes(x = year, y = total)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x= "Academic Semester", y = "Enrollment", title = "USM Enrollment of Hispanic Students")
```
The points on the graph show that each fall semester, the enrollment of Hispanic students across the USM has been increasing, with the exception of FA21, where it decreased from FA20. As usual, this is likely due to the pandemic, however it wasn't that big of decrease. Overall, total enrollment appears to have nearly doubled since FA15, when it was below 12,000, compared to FA24, when it is nearly at 20,000.

```{r}
yearly_totals <- usm_hispanic_enrollment |>
  group_by(year) |>
  summarize(total = sum(enrollment))

model = lm(total ~ year, data = yearly_totals)
summary(model)
```
The model wouldn't run because [according to Google], yearly_totals is a dataset that doesn't exist. I added the first part of the previous code to this code block and it seemed to work. But now the model isn't showing me anything, which, according to Google, is because the years aren't numeric, so I need to fix that. 

I went back to previous homeworks because I remember we did something similar to make values numeric with $, but I can't figure it out. The website GeeksforGeeks is telling me to use as.numeric syntax, however, I don't know how to apply it in the code.


USM HISPANIC ENROLLMENT BY GENDER:
```{r}
usm_hispanic_by_gender |> summarize(correlation = cor(Male, Female, method = "pearson"))
```
```{r}
usm_hispanic_by_gender |> 
  ggplot() + 
  geom_point(aes(x=Male, y = Female)) +
    geom_smooth(aes(x = Male, y = Female), method = "lm", color = "seagreen") +
  labs(title = "USM Male and Female Hispanic Enrollment")
```
No errors with this. Made a scatterplot for fun, but the data and chart show that there is a strong correlation between male and female enrollment. The calculation gave us .967, which is not a perfect correlation but is still very strong. The scatterplot shows this with the line showcasing a positive correlation.

```{r}
usm_hispanic_by_gender |> mutate(male_rate = ("Current Year" - "Prev. Year"/"Prev. Year"))
```
I've tried a couple of things to get the rate change from the previous year's male enrollment but I'm clearly not doing it right. The formula I know is right, but I'm not sure how to do it in R.

---
4/26/2025

```{r}
library(tidyverse)
```

```{r}
usm_hispanic_by_gender <- read_csv("https://raw.githubusercontent.com/emely-miranda/jour405_spring25/refs/heads/main/project/USM_HISPANIC_ENROLLMENT_GENDER.csv")
```
One kind of test I could run using this data on male and female hispanic enrollment from 2015-2024 is measuring the rate of change and if on average, male or female enrollment is growing faster.

I think another test I could run a pearson correlation test to determine whether or not male and female enrollments are moving together. 

```{r}
usm_hispanic_enrollment <- read_csv("https://raw.githubusercontent.com/emely-miranda/jour405_spring25/refs/heads/main/project/USM_HISPANIC_ENROLLMENT.csv")
```
Also, I'm not sure if this is what you meant by longer data but I think I did it right...

With this data I could run similar tests to the above data, such as measuring the rate of enrollment to see if the enrollment of Hispanic students is decreasing or increasing in the different universities.

-----
Prev. assignment
##NOTES:
The table shows the enrollment of Hispanic students across the USM between Fall 2015 and Fall 2024. The data only accounts for fall semester enrollment and not spring.

The most important rows will likely be FA_15 and FA_24, because that will showcase the differences in the number of students enrolling in the colleges.

As I'm looking at the data though, the starting and end points are not far off. If it was a chart/graph, I think there would be flunctuations, but at first glance the data seems constant.

I was wondering if it might be worth exploring the gender aspect of it, but I'm not sure how exactly I would go about that. Are male Hispanic enrollments going down while female Hispanic enrollment goes up, or vice versa. Or are they correlated and both follow the same trend? I would have to make a new table but the data comes from the same place so it will be easy to get. (This idea just came to me so I'm kind of just spitballing here.)


