---
title: "final_project"
output: html_document
---

4/26/2025

```{r}
library(tidyverse)
```

```{r}
usm_hispanic_by_gender <- read_csv("https://raw.githubusercontent.com/emely-miranda/jour405_spring25/refs/heads/main/project/USM_HISPANIC_ENROLLMENT_GENDER.csv")
```
One kind of test I could run using this data on male and female hispanic enrollment from 2015-2024 is measuring the rate of change and if on average, male or female enrollment is growing faster.

I think another test I could run a pearson correlation test to determine whether or not male and female enrollments are moving together. 

```{r}
usm_hispanic_enrollment <- read_csv("https://raw.githubusercontent.com/emely-miranda/jour405_spring25/refs/heads/main/project/USM_HISPANIC_ENROLLMENT.csv")
```
Also, I'm not sure if this is what you meant by longer data but I think I did it right...

With this data I could run similar tests to the above data, such as measuring the rate of enrollment to see if the enrollment of Hispanic students is decreasing or increasing in the different universities.

-----
Prev. assignment
##NOTES:
The table shows the enrollment of Hispanic students across the USM between Fall 2015 and Fall 2024. The data only accounts for fall semester enrollment and not spring.

The most important rows will likely be FA_15 and FA_24, because that will showcase the differences in the number of students enrolling in the colleges.

As I'm looking at the data though, the starting and end points are not far off. If it was a chart/graph, I think there would be flunctuations, but at first glance the data seems constant.

I was wondering if it might be worth exploring the gender aspect of it, but I'm not sure how exactly I would go about that. Are male Hispanic enrollments going down while female Hispanic enrollment goes up, or vice versa. Or are they correlated and both follow the same trend? I would have to make a new table but the data comes from the same place so it will be easy to get. (This idea just came to me so I'm kind of just spitballing here.)


